---
title: "Bayesian linear regression"
author: "Steinunn Gróa Sigurðardóttir"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(blm)
library(MASS)
library(graphics)
```

### A Bayesian Linear Regression package

Bayesian linear regression is a way of doing linear regression, were bayesian statistics are used to do the prediction.

This package is built in RStudio to do Bayesian linear regression, as a part of the course "Data Science in Bioinformatics: Software Development and Testing", in Aarhus University 2016. The goal of the project was to see the various aspects that go into building an R package. 

### Installation and usage
First, install the devtools package. Then install the blm package from Github: install_github("steinunngroa/blm").

### Functions
There are 10 functions available for users of the package. To get information on each function and the usage, type into Rstudio: ?function_name.blm, for example ?deviance.blm. .

- blm()
  The function fits a blm model, using a model formula, and optional additional data.
- coef() 
	This function returns the fitted parameters of the model.
- confint()
	This function finds the confidence intervals for the blm fitted parameters.
- fitted()
  This function produces fitted response variables.
- residuals()
	This function returns the residuals of a blm.
- deviance()
  This function finds the deviance for a blm.
- plot()
  This function produces a plot for a given blm.
- predict()
	This function makes a prediction based on the fitted model.
- print() 
  This function prints out a summary of a blm model
- summary()
  This function gives a summary of a blm regression model, using the print.blm function.


Then there are three support functions used. One to make a prior, another to update to a posterior and the third to give a responseless model matrix. 

- make_prior()
	This function produces a prior distribution, using the parameter alpha and the model given as input.
- update()
	This function updates a prior distribution, to get the posterior distribution.
- model_matrix_responseless()
  This function removes the response variable from a matrix and returns a responseless matrix.
  
### Example of usage

```{r}
# Making of random data for the test
alpha = 1; beta = 1;
w0 = 0.4; w1 = 0.2; w2 = 0.1;
d1 = data.frame(x=rnorm(100), z = rnorm(100));
d1$y = rnorm(100, w0-w1*d1$x+w2*d1$z, 1/beta)
m1 = y ~ x + z

# Using make_prior and blm to construct a valid blm
prior1 = make_prior(m1, alpha)
blm1 = blm(m1, prior1, beta, d1)

# Using coef, confint and fitted
coef1 = coef(blm1)
confint1 = confint(blm1, c("(Intercept)", "x", "z"), level = 0.9)
fit1 = fitted(blm1)

# Showing usage of residuals, deviance and predict
res1 = residuals(blm1)
dev1 = deviance(blm1)
pred1 = predict(blm1, d1)

# Printing, plotting and doing a summary
print(blm1)
plot(blm1)
summary(blm1)
```

## Reference

Mailund, T. "Introduction to Data Science and Statistical Programming in R"